Evolution of CSMA Protocols of the IEEE 802.11 Standard

### 2013-01-22 paper.tex
finished related work (?)

This third group of solutions uses a deterministic backoff after successful transmissions to further reduce the chances of collisions.
It was introduced in \cite{barcelo2008lba} and a more detailed analysis that includes both saturated and non-saturated conditions is presented in \cite{bellalta2009vtc}.
analyzed in depth in \cite{he2009srb}.

>>>>

This third group of solutions uses a deterministic backoff after successful transmissions to further reduce the chances of collisions.
Under certain conditions, collision-free operation is reached.
It was introduced in \cite{barcelo2008lba} and a more detailed analysis that includes both saturated and non-saturated conditions is presented in \cite{bellalta2009vtc}.
A more in-depth study is carried out in \cite{he2009srb}, including realistic elements such as clear channel assessment errors.
Different aspects of fairness are addressed in \cite{he2009srb,barcelo2010fcc,fang2011dlm}.
The performance in realistic channels taking into account the Auto Rate Fallback mechanism is evaluated \cite{martorell2012pec,martorell2012tfl}.

Even though initial research efforts where focused on the WLAN collision problem, more recent papers try to extend the idea to multi-hop networks.
In \cite{hui2011epp}, the multi-hop slotted case explored.
The more realistic situation in multi-hop networks in which the time is not slotted is studied in \cite{barcelo2013dcc}.

The same principles that we exploit to prevent collisions in WLANs can be used in other areas of radio resource management in wireless area networks \cite{duffy2011dcs,checco2012scs,checco2012lbc,khan2013aso}.

In all the work cited so far, there is a limitation that the number of contenders should not exceed the value of the deterministic backoff used after successful transmissions.
If this value is exceeded, it is no longer possible to achieve collision-free operation.
A first solution to solve this problem is presented in \cite{barcelo2011tcf}, but it requires the presence of a central entity (typically the access point) that instructs the other nodes to adjust the value of their deterministic backoff.

In the present paper, we study a completely distributed solution to accommodate a large number of contenders in a fair collision-free fashion.
